
% !TeX root = ../main.tex
\begin{englishabstract}
    % \footnotetext{*The work was supported by the Foundation (foundation ID).}

    Human activity recognition based on Wi-Fi Channel State Information (CSI) has become an important research direction in intelligent sensing due to its non-contact nature, low cost, and privacy-preserving property. CSI captures wireless channel variations caused by human activities and provides an effective basis for indoor action recognition. However, most existing methods assume that all categories are available during training, which makes them difficult to apply in practical scenarios involving continuous category expansion, inaccessible old-class data, and limited samples for newly introduced classes. To address these challenges, this thesis focuses on the continual learning requirements in Wi-Fi CSI-based indoor action recognition and investigates two representative settings, namely class-incremental learning with relatively sufficient new-class samples and few-shot class-incremental learning with extremely limited new-class samples.

    For the class-incremental learning scenario with relatively sufficient new-class samples, this thesis proposes a Replay-Expansion-Consolidation framework. Based on memory replay, representative old-class samples are preserved to alleviate the distribution imbalance between old and new classes during incremental learning. On this basis, the combination of backbone freezing and classifier expansion enables the model to incorporate novel categories while maintaining the stability of old-class representations. Furthermore, a teacher-student structure with bilateral knowledge distillation is introduced to consolidate old and new knowledge into a unified model. For the few-shot class-incremental learning scenario with extremely limited new-class samples, this thesis proposes a learning scheme based on temporal enhancement and graph attention. Specifically, a multi-scale temporal encoder is designed to model action patterns from CSI signals at different temporal scales, a temporal attention fusion module is introduced to aggregate salient information, and an enhanced graph attention network is employed during incremental sessions to capture the relations among class prototypes. In addition, a dual-classifier architecture is adopted to fuse discriminative information from different sources, thereby alleviating bias between old and new classes as well as class-level overfitting under few-shot conditions. Experimental results on the XRF dataset show that the proposed class-incremental learning method maintains favorable overall recognition performance under multi-stage category expansion, achieving 73.57\% accuracy at the final incremental stage. Under the 5-way 1-shot setting, the proposed few-shot class-incremental learning method also demonstrates good stability, achieving 74.24\% accuracy at the final session and an average accuracy of 78.48\%, with overall performance superior to the comparison methods.

    In summary, this thesis addresses catastrophic forgetting and sample scarcity in Wi-Fi CSI-based indoor action recognition under continuous category expansion, and proposes corresponding methods from the perspectives of class-incremental learning and few-shot class-incremental learning. Experimental results verify the effectiveness of the proposed methods. The study provides a feasible solution for the continual updating of Wi-Fi sensing systems in dynamic scenarios and offers a basis for future research on wireless sensing with stronger robustness and generalization ability.

    \englishkeywordstype{Wi-Fi; Indoor Action Recognition; Class-Incremental Learning; Few-Shot Class-Incremental Learning; Knowledge Distillation}{Application Research}

\end{englishabstract}
